Models

### 2024-05-13 analysis/05_model_metrics.Rmd
Move sensitivy analysis into main analysis script so p-values can be adjusted.

####Data exploration

Visualize number of years for each timeseries
```{r}
hist_numYears <- model_data %>%
  select(year, rarefyID) %>% 
  distinct() %>% 
  count(rarefyID, name = "num_years") %>% 
  ggplot() +
  geom_histogram(aes(x = num_years),
                 binwidth = 2) +
  scale_x_continuous(name = 'Number of years sampled',
                     breaks = c(2,4,8,16,32,48,64,96)) +
  #scale_y_continuous(breaks = c(0,2500,5000,7500,10000,12500, 15000, 17500, 20000,22500)) +
  labs(tag = 'B',
       y = 'Number of cells') +
  theme_bw() +
  theme(panel.grid.minor.x = element_blank())

hist_duration <- model_data %>%
  select(rarefyID, duration) %>% 
  distinct() %>% 
  ggplot() +
  geom_histogram(aes(x = duration),
                 binwidth = 2) +
  scale_x_continuous(name = 'Number of years sampled',
                     breaks = c(2,4,8,16,32,48,64,96)) +
  #scale_y_continuous(breaks = c(0,2500,5000,7500,10000,12500, 15000, 17500, 20000,22500)) +
  labs(tag = 'B',
       y = 'Number of cells') +
  theme_bw() +
  theme(panel.grid.minor.x = element_blank())
```

@@

indv_mod_table <- indv_mod_results %>% select(-model)
usethis::use_data(indv_mod_table)
```
Perform p-value correction
```{r}
# Let's get a dataframe of p-values for the models we actually used

@@

pval_df <- bind_rows(base, contrast, group_slope, study_trend) %>%
  filter(metric != "Jaccard_next")

@@

usethis::use_data(model_output)
```

@@

Summaries of study slopes
```{r}
#First - how many studies had a significant trend for each metric? 
model_output %>% 
  filter(!is.na(study_id)) %>%
  filter(term == "year_scaled", p.adjust < 0.05) %>% 
  filter(metric %in% c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_base", "S", "FRic", "FEve", "FDiv")) %>% 
  mutate(positive = ifelse(estimate > 0, TRUE, FALSE)) %>%
  group_by(metric, positive) %>% 
  summarise(n_study_id = n_distinct(study_id)) %>%
  pivot_wider(names_from = "metric", values_from = "n_study_id")

#And which studies were they?
model_output %>% 
  filter(!is.na(study_id)) %>%
  filter(term == "year_scaled", p.adjust < 0.05) %>% 
  filter(metric %in% c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_base", "S", "FRic", "FEve", "FDiv")) %>% 
  select(study_id, metric) %>%
  distinct()

#Which were significant for multiple metrics?
model_output %>% 
  filter(!is.na(study_id)) %>%
  filter(term == "year_scaled", p.adjust < 0.05) %>% 
  filter(metric %in% c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_base", "S")) %>% 
  group_by(study_id) %>%
  filter(n() > 1) %>%
  select(study_id, metric)
```

@@

>>>>

@@

indv_mod_table <- indv_mod_results %>% select(-model)
usethis::use_data(indv_mod_table)
```

#### Sensitivity analysis for duration and trait coverage

1. Min time series duration
```{r}
dur_dfs <- list(min3 = filter(model_data, duration > 2),
                min4 = filter(model_data, duration > 3),
                min5 = filter(model_data, duration > 4))

duration_sens <- map_dfr(dur_dfs, ~pmap_dfr(filter(mods, model_id == "base"), get_model, data = .x), .id = "dataID")

# first group of bad models, try with REML off
bad_dur_models <- duration_sens %>%
  select(metric, model_id, warning, dataID) %>% 
  filter(!is.na(warning)) %>% 
  distinct()

min3_bad <- bad_dur_models %>% filter(dataID == "min3") %>% pull(metric)
min4_bad <- bad_dur_models %>% filter(dataID == "min4") %>% pull(metric)
min5_bad <- bad_dur_models %>% filter(dataID == "min5") %>% pull(metric)

refit_dur <- bind_rows(pmap_dfr(filter(mods, model_id == "base", metric_name %in% min3_bad), get_model, data = dur_dfs$min3, REML = FALSE) %>% mutate(dataID = "min3"),
                       pmap_dfr(filter(mods, model_id == "base", metric_name %in% min4_bad), get_model, data = dur_dfs$min4, REML = FALSE) %>% mutate(dataID = "min4"),
                       pmap_dfr(filter(mods, model_id == "base", metric_name %in% min5_bad), get_model, data = dur_dfs$min5, REML = FALSE) %>% mutate(dataID = "min5")
)

# two left that didn't converge
bad_dur_models2 <- refit_dur %>%
  select(metric, model_id, warning, dataID) %>% 
  filter(!is.na(warning)) %>% 
  distinct()

refit_dur_2 <- bind_rows(pmap_dfr(filter(mods, model_id == "base", metric_name == "CWM_diet_planto"), get_model, data = dur_dfs$min3, bobyqa = TRUE) %>% mutate(dataID = "min3"),
                       pmap_dfr(filter(mods, model_id == "base", metric_name == "CWM_forstrat_midhigh"), get_model, data = dur_dfs$min5, bobyqa = TRUE) %>% mutate(dataID = "min5"))

duration_sens_mods <- bind_rows(duration_sens %>% filter(is.na(warning)),
                                refit_dur %>% filter(is.na(warning)),
                                refit_dur_2 %>% filter(is.na(warning)))
```

2. Trait Coverage
```{r}
load(here::here("data/study_table.rda"))

model_data <- model_data %>%
  left_join(study_table %>% select(study_id, coverage) %>% mutate(study_id = as.factor(study_id)))

covg_dfs <- list(covg75 = filter(model_data, taxa != "Amphibians"),
                covg85 = filter(model_data, taxa != "Amphibians", coverage >= 0.85),
                covg90 = filter(model_data, taxa != "Amphibians", coverage >= 0.90))

coverage_sens <- map_dfr(covg_dfs, ~pmap_dfr(filter(mods, model_id == "base"), 
                                             get_model, data = .x), .id = "dataID")

# first group of bad models, try with REML off
bad_cov_models <- coverage_sens %>%
  select(metric, model_id, warning, dataID) %>% 
  filter(!is.na(warning)) %>% 
  distinct()

covg75_bad <- bad_cov_models %>% filter(dataID == "covg75") %>% pull(metric)
covg85_bad <- bad_cov_models %>% filter(dataID == "covg85") %>% pull(metric)
covg90_bad <- bad_cov_models %>% filter(dataID == "covg90") %>% pull(metric)

refit_cov <- bind_rows(pmap_dfr(filter(mods, model_id == "base", metric_name %in% covg75_bad), 
                                get_model, data = covg_dfs$covg75, REML = FALSE) %>% mutate(dataID = "covg75"),
                       pmap_dfr(filter(mods, model_id == "base", metric_name %in% covg85_bad), 
                                get_model, data = covg_dfs$covg85, REML = FALSE) %>% mutate(dataID = "covg85"),
                       pmap_dfr(filter(mods, model_id == "base", metric_name %in% covg90_bad), 
                                get_model, data = covg_dfs$covg90, REML = FALSE) %>% mutate(dataID = "covg90")
)

coverage_sens_mods <- bind_rows(coverage_sens %>% filter(is.na(warning)),
                                refit_cov %>% filter(is.na(warning)))

sens_mods <- bind_rows(duration_sens_mods, coverage_sens_mods) %>%
  select(-model, -warning) %>%
  filter(metric != "Jaccard_next")
```

#### Perform p-value correction
```{r}
# Let's get a dataframe of p-values for the models we actually used

@@

# sensitivity models
sens_trends <- sens_mods %>%
  filter(term == "year_scaled")

pval_df <- bind_rows(base, contrast, group_slope, study_trend, sens_trends) %>%
  filter(metric != "Jaccard_next")

@@

sensitivity_output <- sens_mods %>%
  select(-p.value) %>%
  left_join(pval_df %>% select(p.value, p.adjust, effect, term, model_id, dataID, metric), na_matches = "never") %>%
  rename(sens_level = dataID)

usethis::use_data(model_output)
usethis::use_data(sensitivity_output)

write_csv(sensitivity_output, here::here("paper", "sensitivity_models.csv"))
```

@@

@@

study_time_data <- model_data %>% 
  dplyr::select(study_id, startyear, duration) %>% 
  distinct() %>%
  group_by(study_id) %>%
  filter(startyear == min(startyear))
